### 2021-12-07 products/manuscript/Supplementary_Data.Rmd

---
title: "Supplementary Data"
author: "C. Coleman"
date: "11/12/2021"
output:
  word_document: default
  html_document: default
---

#### The following markdown serves as the supplementary materials for the UOWN Stream Health Indicators Project. Included materials are all figures relevant to my manuscript and were derived from the analysis code.

```{r,echo=FALSE, results='hide', message=FALSE}
library(knitr)    # For knitting document and include_graphics function
library(ggplot2)  # For plotting
library(png)      # For grabbing the dimensions of png files
```

_Supplementary Figure 4: Cook Distance plot for MIDO Biological Score data. Plots show no score >1, suggesting no outliers in the data._
```{r, echo=FALSE, message=FALSE}
MIDO_BS_OUT_location <- here::here("results", "MIDO_BS_OUT.png")

include_graphics(MIDO_BS_OUT_location)
```

_Supplementary Figure 5: Cook Distance plot for MIDO E. coli cfu data. Plots show no score >1, suggesting no outliers in the data._
```{r, echo=FALSE, message=FALSE}
MIDO_EC_OUT_location <- here::here("results", "MIDO_EC_OUT.png")

include_graphics(MIDO_EC_OUT_location)
```

_Supplementary Figure 6: Cook Distance plot for NORO Biological Score data. Only plot showing score >1 is biological score as a function of Nitrate (NO3) concentration. All other plots have a score <1, suggesting no outliers in the data._
```{r, echo=FALSE, message=FALSE}
NORO_BS_OUT_location <- here::here("results", "NORO_BS_OUT.png")

include_graphics(NORO_BS_OUT_location)
```

_Supplementary Figure 7: Cook Distance plot for NORO E. coli cfu data. Only plot showing score >1 is E. coli cfu as a function of Nitrate (NO3) concentration. All other plots have a score <1, suggesting no outliers in the data._
```{r, echo=FALSE, message=FALSE}
NORO_EC_OUT_location <- here::here("results", "NORO_EC_OUT.png")

include_graphics(NORO_EC_OUT_location)
```

_Supplementary Figure 8: Cook Distance plot for BICO Biological Score data. Plots show no score >1, suggesting no outliers in the data._
```{r, echo=FALSE, message=FALSE}
BICO_BS_OUT_location <- here::here("results", "BICO_BS_OUT.png")

include_graphics(BICO_BS_OUT_location)
```

_Supplementary Figure 9: Cook Distance plot for BICO E. coli cfu data. Only plot showing score >1 is E. coli cfu as a function of Turbidity. All other plots have a score <1, suggesting no outliers in the data._
```{r, echo=FALSE, message=FALSE}
BICO_EC_OUT_location <- here::here("results", "BICO_EC_OUT.png")

include_graphics(BICO_EC_OUT_location)
```